---
title: "Regression Models Course Project"
author: "Victor Bucha"
date: "21th of October"
output: pdf_document
---

#Executive summary
In this paper we explored the relationship between miles-per-gallon (MPG) and two other variables. First variable is a trasmission type and second one is a car weight. The last variable has the highest correlation with MPG. 

At exploatory and inference section we have shown that cars with manual transmission are better in terms of MPG. 

Later we studied relationship of MPG and other variables by building several regression models (MPG ~ Weight+Transmission) and (MPG ~ Weight*Transmission). 

Although the best linear regression model allows to predict ~83% of MPG variability, we can not use it to judge what transmission is better in respect to weight and transmission type.The main reason is that mtcars dataset have very small overlap between the trasmission groups for any given weight. 

#Exploatory analysis 
At first, let us create boxplot showing the distribution of mpg for different types of trasmission. It looks like cars with automatic transmisson tend to have lower mpg.
```{r, Exploratory analysis, fig.width=6, fig.height=3.2}
data(mtcars) #load data
library(ggplot2)
p <- ggplot(mtcars, aes(factor(am), mpg))
p + geom_boxplot(aes(fill=factor(am)))+
        labs(title="MPG vs Transmission",x="Transimssion type",y="MPG")+
        scale_fill_discrete("Transmission",labels=c("Auto", "Manual"))
```

## Inference
Assuming the data has a normal distribution let us make t-test in order to show that automatic transmission has lower mean mpg value than cars with manual transmission. We will test following hypotheses H0: mu=0, H1: mu<0.
```{r, T-test}
t.test(subset(mtcars,am==0,mpg),subset(mtcars,am==1,mpg),alternative="less")
```
P-value is small and both confidence interval values are negative. So we reject null hypothesis and accept alternative. It means cars with automatic transmission has less average MPG than cars with manual transmission with 95% confidence.

#Regression analysis
In this section, we start to build several regression models. We started from simple linear regression model with one regressor. Then proceed with multivariable regression model by addding the variable that has the strongest correlation with mpg. Finnaly, we build multivariable regression model that have the best prediction strength.

##Base regression model
Let us build simple linear reggression model with [mpg] as outcome and one regressor as [am].
```{r, Base regression model}
m1<-lm(mpg~am,mtcars)
summary(m1)
```
### Coefficients interpretation
* Beta0/intercept is equal to 17.147. It corresponds to the mean mpg value for cars with automatic transmission
* Beta1 is equal to 7.245. It increases of MPG for cars with manual transmission.
* Adjusted R-squared is equal 0.3385. So ~34% of mpg variability can be explained by transmission type. 

## Multivariable regression 
In previous section, we saw that the base regression model is very weak in prediction. Let us find the predictor whose correlation with MPG is the highest.

```{r, Compute correlation}
cor(mtcars)[,1]
```

Weight [wt] has the highest negative correlation -0.8676594 with MPG. Let us select it as a second predictor and build new linear regression model.

```{r, Multivariable regression model}
m2<-lm(mpg~wt+factor(am),mtcars)
summary(m2)
```
### Coefficients interpretation
* Adjusted R-squared grows more than two times. With a new model we can explain ~75% of MPG variability. 
* It's interesting that [am] slope coefficient beta2 changed the sign and p-value become extremely large 0.988. That means beta2 is not statistically important in this model.

Transmission type [am] is a binary variable taking 0 for auto and 1 for manual transmissions. Above model fits parallel lines for the two levels of factor variable [am]. Since [am] slope coefficient beta2 is close to zero (-0.02362) it adds a litle to intercept of the second parallel line. So two lines are almost identical. Bellow figure show it in more details.

```{r, Model plot, echo=FALSE, fig.width=6, fig.height=3.2}
ggplot(mtcars,aes(y=mpg,x=wt))+
        geom_point(aes(colour=factor(am)),size=8,alpha=0.7)+
        geom_point(colour="grey90",size=5,alpha=0.6)+
        geom_abline(intercept=m2$coefficients[1],slope=m2$coefficients[2],
                    aes(colour="0"),linetype=1,size=2)+
        geom_abline(intercept=m2$coefficients[1]+m2$coefficients[3],
                    slope=m2$coefficients[2],aes(colour="1"),linetype=2,size=2)+
        labs(title="lm(mpg~wt+factor(am),mtcars)",x="Weight",y="MPG")+
        scale_color_discrete("Transmission",labels=c("Auto", "Manual"))
```

## Multivariable regression with interaction
Let us make interaction with variable [am] and [wt] in order to improve our model. 
```{r, Multivariable regression model 2}
m3<-lm(mpg~wt*factor(am),mtcars)
summary(m3)
```
### Coefficients interpretation
* Adjusted R-squared grows up to 0.833. With a new model we can explain ~83% of MPG variability.
* P-values are small. So all coefficients are statisticaly important

Now our model with inteaction term allows both a diffirent intercept and slope for each group. See bellow figure for more details. Because of that we have better prediction model.

```{r, Model interaction plot, echo=FALSE}
ggplot(mtcars,aes(y=mpg,x=wt))+
        geom_point(aes(colour=factor(am)),size=8,alpha=0.7)+
        geom_point(colour="grey90",size=5,alpha=0.6)+
        geom_abline(intercept=m3$coefficients[1],slope=m3$coefficients[2],
                    aes(colour="0"),linetype=1,size=2)+
        geom_abline(intercept=m3$coefficients[1]+m3$coefficients[3],
                    slope=m3$coefficients[2]+m3$coefficients[4],
                    aes(colour="1"),linetype=2,size=2)+
        labs(title="lm(mpg~wt*factor(am),mtcars)",x="Weight",y="MPG")+
        scale_color_discrete("Transmission",labels=c("Auto", "Manual"))
```

##Selection of the best model
Let us select the best model by using anova to test the models against one another.
```{r, Best model selection}
anova(m1,m2,m3)
```
As we can see, second model: m2=lm(mpg~wt+am) is better than the first one: m1=lm(mpg~am) because anova test show positive difference of RSSs= 442.58 and P-value for F statistic is small= 7.717e-09.

However third model m3=lm(mpg~wt*am) with interaction term is better than second one m2=lm(mpg~wt+am) since third model further reduce sum of squared differences and p-value is still low=0.001017.

So the third model m3=lm(mpg~wt*am) is selected to be the best one.

##Residuals and Diagnostics
Let us look to residuals of the model m3=lm(mpg~wt*am) that we selected as the best one.
```{r, Risidual for best model}
par(mfrow = c(2, 2))
plot(m3)
```
From the above plot, we can make following observation:  

* The points in the Risidual vs. fitted plot seem to have linear negative correlation which may violate independence condition.  
* Normal Q-Q plot has points that do not fall on the line indicating that risiduals might be not normally distributed  
* The Scale-Location plot shows points in constan band, indicating constant variance  

#Conclusion
Although our third model was selected as the best in terms of MPG prediction it can be controversial in making statement what trasmission is better. Our regression model says that automatic transmission is better in terms of MPG for cars with a weight greater than 3 tones. For cars with weight less than 3 tons automatic tranmission is worse. However, we can not rely on above statements because of following reasons:

At first, weight variable [wt] is highly related to transmission type [am]. Take a look at bellow plot. Almost all cars that have automatic transmission have the weight greather than 3 tons. Also in this setting, there is small overlap between the trasmission groups for any given weight. That means there isn't a lot of evidence to compare the groups. Above analysis of risiduals supports these observations. 

```{r, Model interaction plot 2, echo=FALSE}
ggplot(mtcars,aes(y=mpg,x=wt))+
        geom_point(aes(colour=factor(am)),size=8,alpha=0.7)+
        geom_point(colour="grey90",size=5,alpha=0.6)+
        geom_abline(intercept=m3$coefficients[1],slope=m3$coefficients[2],
                    aes(colour="0"),linetype=1,size=2)+
        geom_abline(intercept=m3$coefficients[1]+m3$coefficients[3],
                    slope=m3$coefficients[2]+m3$coefficients[4],
                    aes(colour="1"),linetype=2,size=2)+
        labs(title="lm(mpg~wt*factor(am),mtcars)",x="Weight",y="MPG")+
        scale_color_discrete("Transmission",labels=c("Auto", "Manual"))
```
