---
title: "Regression Models Course Project"
author: "Victor Bucha"
date: "19th of October"
output: html_document
---
# Task
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

Q1: Is an automatic or manual transmission better for MPG  
Q2: Quantify the MPG difference between automatic and manual transmissions

Your report must be:
Written as a PDF printout of a compiled (using knitr) R markdown document.
Brief. Roughly the equivalent of 2 pages or less for the main text. Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures.
Include a first paragraph executive summary.

#Executive summary

##Is an automatic or manual transmission better for MPG?

##Quantify the MPG difference between automatic and manual transmissions

#Appendix


```{r, Q1}
data(mtcars) #load data
library(ggplot2)

mtcars$am<-factor(mtcars$am) # convert mtcars$am to factor
levels(mtcars$am)<-factor(c("auto","manual")) # rename factor levels

auto_tr<-mtcars[mtcars$am=="auto",]# select cars with auto transmission
manual_tr<-mtcars[mtcars$am=="manual",]# select cars with auto transmission

#compute mean mpg value for car with auto trasmission
mean_mpg_auto<-mean(auto_tr$mpg) 
#compute mean mpg value for car with manual trasmission
mean_mpg_manual<-mean(manual_tr$mpg) 

#draw plot of MPG vs Transmission
ggplot(mtcars,aes(x=am,y=mpg,shape=am))+
        geom_point(aes(colour=am),size=4)+
        geom_point(colour="grey90", size = 1.5)+
        geom_hline(aes(yintercept=mean_mpg_auto,color="auto"),size=1.1,linetype=2)+
        geom_hline(aes(yintercept=mean_mpg_manual,color="manual"),size=1.1,linetype=2)+
        labs(title="Cmparision MPG for auto and manual transmission", 
             x="Transimssion",y="MPG")+
        annotate("text",label=paste(paste(rep(" ",24),collapse = ""),
                                    paste("Mean:",format(mean_mpg_auto,digits=4))),
                                    y=mean_mpg_auto+0.8,x="auto")+
        annotate("text",label=paste(paste(rep(" ",24),collapse = ""),
                                    paste("Mean:",format(mean_mpg_manual,digits=4))),
                                    y=mean_mpg_manual+0.8,x="manual")+
        scale_color_discrete("Transmission")+scale_shape_discrete("Transmission")
```

Cars with automatic transmission  seems to provide less MPG. Let us check these hypotheses
with t-test as follows H0: mu=0, H1: mu<0

```{r, Q1-t-test}
t.test(auto_tr$mpg,manual_tr$mpg,alternative="less")
```
We reject Null hypothesis and accept alternative. It means cars with automatic transmission has less
average MPG than cars with manual transmission with 95% confidence.

```{r, Q1-additional}

ggplot(mtcars,aes(y=mpg,x=hp))+
        geom_point(aes(colour=factor(am)),size=8,alpha=0.7)+
        geom_point(colour="grey90",size=5,alpha=0.6)

```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```


